---
title: "Data Visualization"
author: "Grace Lewin"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General use
library(tidyverse)
library(here)
library(janitor)

# Specifically for dataviz:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)
```


## Get the Lizards Data

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

Histogram (really bad) of lizard weights

```{r}
ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(fill = "orange", 
                 color = "red",
                 size = 0.2,
                 linetype = "dotted")
```


Scatterplot

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 22,
             color = "green",
             fill = "purple",
             size = 4,
             alpha = 0.5)
```

Bubble plot where the color of the points changes based on common_name and the size of points varies based on total_length.

DISCLAIMER: BAD IDEA!

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, size = total_length), 
             shape = 22,
             fill = "black") +
  theme_minimal() 
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name, ncol = 4)
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_grid(sex ~ tail)
```


Find total lizard counts by common name
```{r}
# This way would make more sense if you're doing more summary variables than just count. Then you could put mean, sd, etc in with n() within the summarize function.
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

# does the same thing in this case. 
lizard_counts <- lizards %>%  dplyr::count(common_name)
lizard_counts

lizard_counts_cn_tail <- lizards %>% count(common_name, tail)
lizard_counts_cn_tail
```


```{r}
# ggplot(data = lizard_counts, aes(x = common_name, y = n)) +
 # geom_col() +
  #coord_flip()
  
# fct_reorder() creates a factor and has it ordered by variable specified
ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, n), 
                                 x = n)) +
  geom_col(aes(fill = common_name), show.legend = FALSE) +
  labs(x = "Lizard Counts", y = "Common Name")
  
```

Try converting common name to an ordered factor outside of ggplot.

```{r}
lizard_counts = lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, n))
```


## Axis scale breads, limits and labels

Scatterplot: total_length (x) vs. weight (y)

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, 10, 50, 500),
                     limits = c(c(0, 500)),
                     expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))
```

Transform the date column to class Date, then find counts of observations by date.
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)
```


Make a line plot (geom_line()) of date (x) and count (y)
```{r}
ggplot(data = lizard_counts, aes(x = date, y = n)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%Y") #%Y specifies 4 digit year, %y would do 2 digit year, etc. 
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_stepsn(colors = c("green", "blue", "purple"), 
                     breaks = c(0, 20, 40, 50, 60))
```


### Update a color scheme using a palette in paletteer
Make a horizontal boxplot with common_name on the y-axis, total_length on the x-axis, with color changing based on common_name.

```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

ggplot(data = lizards_fct, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "calecopal::sage")
```


```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
     panel.background = element_rect(fill = "yellow", 
                                     color = "purple",
                                     size = 10),
    # panel.grid.major.x = element_line(color = "red"),
    # panel.grid.minor.x = element_line(color = "green"),
    # panel.grid.major = element_line(color = "blue"),
    # panel.grid.minor.y = element_line(color = "magenta")
    panel.grid = element_blank(), # gets rid of gridlines.
    axis.text.x = element_text(color = "orange"),
    axis.title = element_text(color = "green", size = 15)
  )
```



